---
title: "Final Infographic"
author: "Luna Herschenfeld-Catalán she/her"
date:  2024-03-09
format:
  html:
    embed-resources: true
output:
  html_document:
    code_folding: hide
---

```{r include = FALSE}

# Read in libraries
library(tidyverse)
library(here)
library(lubridate)
library(janitor)
library(sf)
library(tidycensus)
library(patchwork)

# do google fonts
library(monochromeR)
library(showtext)
library(ggtext)
library(ggrepel)

knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```


# Data

## Wrangle and Clean Data
 https://www.census.gov/cgi-bin/geo/shapefiles/index.php 

```{r message = FALSE, warning = FALSE, results='hide'}
# geometry of the puma 2022 shapefiles
puma <- read_sf(here("data", "tl_2022_06_puma20", "tl_2022_06_puma20.shp"))

puma_la <- puma %>% 
  filter(grepl(037, PUMACE20)) %>% 
  rename(PUMA = PUMACE20)

pums_housing_22 <- read_csv(here::here("data", "pums_2022.csv"))

# https://www.huduser.gov/portal/datasets/il/il2020/2020summary.odn
people <- c(1, 2, 3, 4, 5, 6, 7, 8)
very_low <- c(39450, 45050, 50700, 56300, 60850, 65350, 69850, 74350)
extreme_low <- c(23700, 27050, 30450, 33800, 36550, 39250, 41950, 44650)
low <- c(63100, 72100, 81100, 90100, 97350, 104550, 111750, 118950)

income_breaks <- data.frame(people, extreme_low, very_low, low)

# remove the duplicates from 2022 data
puma_house_codes <- pums_housing_22 %>% 
  select(SERIALNO, PUMA) %>% 
  distinct()

puma_22 <- pums_housing_22 %>% 
  inner_join(puma_la, by = "PUMA") %>% 
  select(SERIALNO, WGTP, NP, HINCP, FINCP, GRPIP) %>% 
  mutate(NP = as.numeric(NP),
         HINCP = as.numeric(HINCP),
         FINCP = as.numeric(FINCP),
         GRPIP = as.numeric(GRPIP)) %>% 
  group_by(SERIALNO) %>% 
  summarize_all(mean, na.rm = TRUE) %>% 
  left_join(puma_house_codes)

la_tract_vacant <- get_acs(
  state = "CA",
  county = "Los Angeles",
  geography = "tract",
  variables = "B25004_001",
  geometry = TRUE,
  year = 2022
)

la_tract <- la_tract_vacant %>%
  mutate(NAME = gsub("; Los Angeles County; California", # elements that you want to remove
                                  "", # replace with blank
                                  NAME)) %>%
  mutate(NAME = gsub("Census Tract ", # elements that you want to remove
                                  "", # replace with blank
                                  NAME)) %>% 
  filter(GEOID != "06037599100") %>% # islands
  filter(GEOID != "06037599000") %>% # islands
  filter(GEOID != "06037980003") %>%
  filter(GEOID != "06037980004") %>%
  filter(!(NAME >= 9000 & NAME <= 9800))
  
```



## Looking at the places where affordable housing projects started

```{r, message=FALSE}
# read in affordable housing projects in LA from 2003 to present
ah_raw <- read_csv(here("data/LAHD_Affordable_Housing_Projects_List__2003_to_Present__20240119.csv")) %>% 
  clean_names()

# clean affordable housing data 
ah_clean <- ah_raw %>% 
  mutate(fun_date = as.Date(date_funded, tryFormats = c("%m/%d/%Y"))) %>% # make date_funded as date
  mutate(year = lubridate::year(fun_date)) %>% # make into year column 
  select(name, year, fun_date, construction_type, site_community, total_units = project_total_units, 
         housing_type, lahd_funded, in_service_date, gps_coords_on_map) %>% 
  mutate(gps_coords_on_map = gsub("[POINT()]", # elements that you want to remove
                                  "", # replace with blank
                                  gps_coords_on_map)) %>% # remove these elements from gps column 
  separate_wider_delim(gps_coords_on_map, 
                       delim = " ", names = c("empty", "coords"), # separate space from before the coordinates
                       too_many = "merge") %>% 
  separate_wider_delim(coords, delim = " ", names = c("Longitude", "Latitude"), # split lat and long coords
                       too_many = "merge") %>% 
  select(-empty) %>% 
  st_as_sf(coords = c("Longitude", "Latitude"), # make into geometry object
                 crs = st_crs(la_tract)) %>% 
  st_join(la_tract) %>% 
  mutate(site_community = str_to_title(site_community))
```

# Assigning income categories to households 

## Create income breaks for households above 8 (from 9-20)
```{r message = FALSE}
# very low income 
size <- seq(9, 20, 1)
very_low <- data.frame()
row_n <- data.frame()

for (i in seq_along(size)) {
  income <- 1.32 + (i*.08)
  limit <- income*56300
  row <- size[i]
  row_n <- rbind(row_n, row)
  very_low = rbind(very_low, limit)
}

very_low <- cbind(row_n, very_low) %>% 
  rename(people = X9,
         very_low = X78820)

extreme <- data.frame()
for (i in seq_along(size)) {
  # https://www.huduser.gov/portal/datasets/il/il2020/2020ILCalc3080.odn#calculator
  income <- 44120 + (i*4480)
  extreme <- rbind(extreme, income)
}

low <- data.frame()
for (i in seq_along(size)) {
  income <- 1.32 + (i*.08)
  limit <- income*90100
  low <- rbind(low, limit)
}

cols_9_20 <- cbind(extreme, very_low) %>% 
  rename(extreme_low = X48600) %>% 
  cbind(low) %>% 
  rename(low = X126140)

income_breaks <- rbind(income_breaks, cols_9_20)
```

## Creating For Loop

### Applying for loop to the whole data set

```{r message = FALSE}
puma_22$status <- rep(NA, nrow(puma_22))

for (person in 1:nrow(puma_22)) {
  
  individual <- puma_22$NP[person]
  
  income <- income_breaks %>% 
    filter(people == individual)
  
  #print(individual)
  #print(individual)
  # Define thresholds for labels
  extreme_threshold <- income[[2]]
  very_low_threshold <- income[[3]]
  low_threshold <- income[[4]]
  
  # # Assign labels based on thresholds
  puma_22$status[person] <- ifelse(puma_22[["HINCP"]][person] <= extreme_threshold, 
                              "extreme",
                              ifelse(puma_22[["HINCP"]][person] > extreme_threshold & puma_22[["HINCP"]][person] <= very_low_threshold, 
                                     "very low",
                                     ifelse(puma_22[["HINCP"]][person] > very_low_threshold & puma_22[["HINCP"]][person] <= low_threshold, 
                                            "low",
                                            ifelse(puma_22[["HINCP"]][person] > low_threshold, "not eligible", "non"))))
}

```


# Mapping census tracts in LA

```{r message = FALSE}
puma_geo <- puma_la %>% 
  select(PUMA, geometry)


puma_plot <- puma_22 %>% 
  group_by(PUMA) %>% 
  count(status) %>% 
  pivot_wider(names_from = "status", values_from = "n") %>% 
  ungroup() %>% 
  clean_names() %>% 
  select(puma, extreme, very_low, low, not_eligible) %>% 
  mutate(PUMA = puma, 
         total_households = rowSums(.[2:5]),
         total_eligible = rowSums(.[2:4]),
         percent_eligible = (total_eligible/total_households) * 100 ) %>% 
  left_join(puma_geo) %>% 
  st_as_sf()

# crop puma plot with the la_tract to remove the top part of LA
puma_plot_crop <- st_crop(puma_plot, la_tract)


# has 2196 points
tract_neighborhood <- read_csv(here("data/Census_Tract_Locations__LA__20240206.csv")) %>% 
  mutate(census_tract = gsub(", Los Angeles County, California", # elements that you want to remove
                                  "", # replace with blank
                                  Tract)) %>% 
  select(-Tract, census_tract, Neighborhood, Latitude, Longitude) %>% 
  st_as_sf(coords = c("Longitude", "Latitude"), # make into geometry object
                 crs = st_crs(puma_plot)) %>% 
  st_crop(puma_plot_crop) # neighborhoods only in the census tracts of interest

# make a df with the puma information and census tract geometry 
puma_tract <- puma_plot_crop %>% # has 67 observations
  st_join(tract_neighborhood,
                join = st_intersects) %>% # has the same number of points as the neighborhood
  clean_names() %>% 
  select(neighborhood, census_tract, location, puma, geometry) %>% 
  mutate(census_tract = gsub("Census Tract ", # elements that you want to remove
                                  "", # replace with blank
                                  census_tract))


# rename the tract column
la_tract <- la_tract %>% 
  rename(census_tract = NAME) %>% 
  select(census_tract, geometry)

puma_tract <- puma_tract %>% 
  st_drop_geometry() %>% 
  left_join(la_tract)

puma_eligibility <- puma_plot %>% 
  st_drop_geometry() %>% 
  select(!PUMA)

# make map of the census tracts with the puma household data
tract_eligibility <- puma_tract %>% 
  left_join(puma_eligibility, by = "puma") %>% 
  filter(!st_is_empty(geometry)) %>% 
  st_as_sf()
```

# Looking at census tract and the percent of income spent on rent: Rent Burdened

```{r}

rent_burden <- puma_22 %>% 
  group_by(PUMA) %>% 
  rename(percent_rent = GRPIP) %>% 
  select(WGTP, percent_rent, PUMA) %>% 
  mutate(status = ifelse(percent_rent >= 30, "burden", "not_burden")) %>% 
  count(status) %>% 
  pivot_wider(names_from = "status", values_from = "n") %>% 
  ungroup() %>% 
  rename(puma = PUMA) %>% 
  mutate(total = burden+not_burden,
         percent_burden = (burden/total)*100)
  
#plot(rent_burden_2$geometry)

rent_burdened <- puma_tract %>% 
  left_join(rent_burden) %>% 
  filter(!st_is_empty(geometry)) %>% 
  st_as_sf()

```


# Creating a custom theme
```{r}
library(monochromeR)
library(showtext)
library(ggtext)
library(ggrepel)

# import google fonts 
font_add_google(name = "Josefin Sans",
                family = "josefin") # name we provide ggplot

font_add_google(name = "Sen",
                family = "sen") # name we provide ggplot

font_add_google(name = "Tenor Sans",
                family = "tenorSans") # name we provide ggplot

# Pairs
font_add_google(name = "Playfair Display",
                family = "playfairDisplay") # name we provide ggplot

font_add_google(name = "Alice",
                family = "alice") # name we provide ggplot

# enable show text here that configures font across platforms
showtext_auto()

my_theme <- theme(
  plot.title = element_text(size = 20, face = "bold", family = "playfairDisplay", hjust = 0.5), # Font size set to 16 and bold.
  plot.caption.position = "plot",
  plot.caption = element_text(size = 10, hjust = 0, family = "alice"),
  axis.title = element_text(size = 14), # Font size set to 12.
  axis.text = element_blank(), # no axis text.
  axis.ticks = element_blank(),
  legend.title = element_text(size = 15, face = "bold", family = "alice"), # Font size of the title of the legend set to 12 and bold.
  legend.text = element_text(size = 10), # font size of the text in the legend
  panel.background = element_rect(fill = "white"), # color of the panel background
  panel.grid.major = element_blank(),
  plot.margin = margin(12, 6, 12, 18),
  panel.grid.minor = element_blank(), # invisible auxiliary grids
  plot.background = element_rect(fill = "white"), # plot's background
  strip.background = element_rect(fill = "white", color = "#ff975d"), # strip background
  strip.text = element_text(size = 12, face = "bold", family = "alice"), # strip texts
  legend.position = "top", # position of the legend
  legend.box.background = element_rect(color = "white"), # background of the plot
  legend.key.size = unit(1, "cm") # size of the legends key
)
```

# Figures 

##
```{r fig.width = 11, fig.height = 20}

my_brew_palette15 <- RColorBrewer::brewer.pal(n = 20, name = 'YlGnBu')
my_brew_palette10 <- RColorBrewer::brewer.pal(n = 10, name = 'RdBu')
my_brew_palette5 <- RColorBrewer::brewer.pal(n = 10, name = 'Reds')

# eligibility of housing in LA
tract_eligibility_percent <- ggplot(tract_eligibility) +
  geom_sf(aes(fill = percent_eligible)) +
  scale_fill_stepsn(colors = my_brew_palette5,
                    labels = scales::label_percent(scale = 1),
                    breaks = scales::breaks_width(width = 7)) +
  labs(title = "Exploring Affordable Housing Needs in Los Angeles County, CA",
       fill = "Percent Eligible",
       caption = "Households Eligible for Affordable Housing by PUMA in Los Angeles, CA") + 
  guides(fill = guide_colorbar(barheight = unit(1, units = "mm"),  
                                 barwidth = unit(100, units = "mm"),
                                 direction = "horizontal",
                                 ticks.colour = "grey20",
                                 title.position = "top",
                                 label.position = "bottom",
                                 title.hjust = 0.5)) +
  my_theme 




rent_burden_plot <- ggplot(rent_burdened) +
  geom_blank() +
  geom_sf(aes(fill = percent_burden)) +
  scale_fill_stepsn(colors = my_brew_palette5,
                    breaks = scales::breaks_width(width = 10),
                    labels = scales::label_percent(scale = 1)) +
  labs(title = "Exploring Affordable Housing Needs in Los Angeles County, CA",
       fill = "Percent Burdened",
       caption = "Percent of Households Rent Burdened by Census Tract in Los Angeles, CA (2022). Dots represent location of affordable housing projects from 2003-2020.") +
  guides(fill = guide_colorbar(barheight = unit(1, units = "mm"),  
                                 barwidth = unit(100, units = "mm"),
                                 direction = "horizontal",
                                 ticks.colour = "grey20",
                                 title.position = "top",
                                 label.position = "bottom",
                                 title.hjust = 0.5)) +
  my_theme


burden_affordable_projects <- rent_burden_plot +
  geom_sf(data = ah_clean, alpha = 0.5, size = 1)

tract_eligibility_percent
burden_affordable_projects

```


# Selecting the Census Tracts of interest
```{r}

# joining the eligibility and rent data together by census tract
burden_eligible <- rent_burdened %>% 
  st_drop_geometry() %>% 
  left_join(tract_eligibility, by = "census_tract") %>% 
  st_as_sf()

top_burden <- burden_eligible %>% 
  slice_max(order_by = percent_burden,
            n = 75)

# high burden, no affordable housing
burden_non <- top_burden %>% 
  slice_min(order_by = percent_burden, 
            n = 25) 

# high burden, affordable housing
burden_ah <- burden_eligible %>% 
  slice_max(order_by = percent_burden,
            n = 30)

plot(burden_ah$geometry)

# low burden, with affordable housing - and also have high percentage eligible for affordable housing
lowburden_ah <- burden_eligible %>% 
  slice_max(order_by = percent_eligible,
            n = 450) %>% 
  slice_min(order_by = percent_eligible,
            n = 25)

plot(lowburden_ah$geometry)

# low burden, no affordable housing - low percentage of households that qualify for affordable housing
lowburden_non <- burden_eligible %>% 
  slice_min(order_by = percent_eligible,
            n = 100) %>% 
  slice_min(order_by = percent_eligible,
            n = 20)

plot(lowburden_non$geometry)

# neighborhoods right next to undeserved geometry represented by `burden_non`
interest <- burden_eligible %>% 
  slice_min(order_by = percent_burden,
            n = 600) %>% 
  slice_max(order_by = percent_burden,
            n = 15) 

plot(interest$geometry)
# %>% 
#   st_union()

```

# Plotting the three comparison neighborhoods 

```{r fig.width = 11, fig.height = 11}

# high burden, no affordable housing
burden_non_plot <- burden_non %>% 
  st_union() %>% 
  ggplot() +
  geom_sf() +
  labs(title = "Neighborhoods with High Rent Burden, \nNo Affordable Housing") +
  my_theme

# high burden, affordable housing
burden_ah_plot <- burden_ah %>% 
  st_union() %>% 
  ggplot() +
  geom_sf() +
  labs(title = "Neighborhoods with High Rent Burden, \nWith Affordale Housing") +
  my_theme

# low burden, with affordable housing - and also have high percentage eligible for affordable housing
lowburden_ah_plot <- lowburden_ah %>% 
  st_union() %>% 
  ggplot() +
  geom_sf() +
  labs(title = "Neighborhoods with Low Rent Burden, \nWith Affordale Housing",
       subtitle = "and also have high percentage eligible for affordable housing") +
  my_theme

# ow percentage of households that qualify for affordable housing
lowburden_non_plot <- lowburden_non %>% 
  st_union() %>% 
  ggplot() +
  geom_sf() +
  labs(title = "Neighborhoods with Low Rent Burden, \nNo Affordale Housing",
       subtitle = "low percentage of households that qualify for affordable housing") +
  my_theme

interest_plot <- interest %>% 
  st_union() %>% 
  ggplot() +
  geom_sf() +
  labs(title = "Neighborhood with low rent burden right next \nto neighborhood with high rent burden and no support",
       subtitle = "50% percentage of households that qualify for affordable housing, less than 20% rent burden") +
  my_theme

burden_non_plot / burden_ah_plot / lowburden_ah_plot / lowburden_non_plot / 
interest_plot

```

```{r fig.width = 11, fig.height = 11}

base_map <- ggplot(rent_burdened) +
  geom_blank() +
  geom_sf(aes(fill = percent_burden)) +
  scale_fill_stepsn(colors = my_brew_palette5,
                    breaks = scales::breaks_width(width = 10),
                    labels = scales::label_percent(scale = 1)) +
  labs(title = "Exploring Affordable Housing Needs in Los Angeles County, CA",
       fill = "Percent Burdened") +
  guides(fill = guide_colorbar(barheight = unit(1, units = "mm"),  
                                 barwidth = unit(100, units = "mm"),
                                 direction = "horizontal",
                                 ticks.colour = "grey20",
                                 title.position = "top",
                                 label.position = "bottom",
                                 title.hjust = 0.5)) +
  my_theme +
  geom_sf(data = ah_clean, alpha = 0.5, size = 1)

# burden_non points
burden_non_pt <- data.frame(
  y = c(33.9884335),
  x = c(-118.1758795),
  yend = c(33.77222),
  xend = c(-117.92694)
)

# burden_ah points
burden_ah_pt <- data.frame(
  y = c(34.0597395),
  x = c(-118.2982),
  yend = c(33.62815),
  xend = c(- 118.30718)
)

lowburden_ah_pt <- data.frame(
  y = c(34.245882),
  x = c(-118.4209225),
  yend = c(33.93291),
  xend = c(-118.66136)
)

lowburden_non_pt <- data.frame(
  y = c(33.860177),
  x = c(-118.388831),
  yend = c(33.82235),
  xend = c(-118.53424)
)
 
# add the lines to the plot in segments - still editing this 
base_map +
geom_segment(data = burden_non_pt,
               aes(x = x, y = y, xend = xend, yend = yend),
               linetype = "dashed",
             color = "black") +
  geom_segment(data = burden_ah_pt,
               aes(x = x, y = y, xend = xend, yend = yend),
               linetype = "dashed",
             color = "black") +
  geom_segment(data = lowburden_ah_pt,
               aes(x = x, y = y, xend = xend, yend = yend),
               linetype = "dashed",
             color = "black") +
  geom_segment(data = lowburden_non_pt,
               aes(x = x, y = y, xend = xend, yend = yend),
               linetype = "dashed",
             color = "black") +
  inset_element(interest_plot, -118.17398, -118.16851, -118.14058, -118.13931, 
                on_top = TRUE)
```


```{r fig.width = 8, fig.height = 5}

RColorBrewer::brewer.pal(n=10,"Reds")

# looking at the total construction of affordable housing in each year 
lahd_funding <- ah_clean %>% 
  drop_na() %>% 
  group_by(year) %>% 
  summarize(total_funding = sum(lahd_funded)) %>% 
  ggplot(aes(x = year, y = total_funding)) +
  geom_col(fill = "#67000D") +
  scale_y_continuous(expand=c(0,0),
                     labels = function(x) paste0(x / 1e6, "M")) +
  theme_classic() +
  labs(x = 'Year',
       y = 'Total Funding (in millions)')

lahd_funding
```


# Next steps 

- get information for these areas 
- Number of households (approximate)
- How has that changed since 2003 

# Look at the industry / jobs of people in the census tract 

This data does not exist
```{r}
# load list of variables that I could access
variables <- load_variables(2022, "acs1", cache = TRUE)

#B24114 is for Detailed Occupation for the Civilian Employed Population 16 Years and Over
#B24115 is for Detailed Occupation for the Civilian Employed Population 16 Years and Over - Male
#B24116 is for Detailed Occupation for the Civilian Employed Population 16 Years and Over - Female
B24114 <- variables %>% 
  filter(str_detect(name, "B24114"))

#B24114$name

la_tract_occupation <- get_acs(
  state = "CA",
  county = "Los Angeles",
  geography = "state",
  variables = "B24114_100",
  geometry = TRUE,
  year = 2021
)

v <- data(acs5_geography)
```

## Try this that looks at the number of workers in household based on size
```{r}
# B08202 Household Size by Number of Workers in Household

B08202 <- variables %>% 
  filter(str_detect(name, "B08202"))

la_tract_workers <- get_acs(
  state = "CA",
  county = "Los Angeles",
  geography = "state",
  variables = B08202$name,
  geometry = TRUE,
  year = 2022
)
```



# Look at the average age / demographic data by tract
https://walker-data.com/tidycensus/articles/basic-usage.html

```{r}
# B01001A is Sex by Age (White Alone)
# B01001B is Sex by Age (Black or African American Alone)
# B01001

 B01001 <- variables %>% 
  filter(str_detect(name, "B01001"))

# has the de
la_tract_age <- get_acs(
  state = "CA",
  county = "Los Angeles",
  geography = "tract",
  variables = B01001$name,
  geometry = TRUE,
  year = 2022
)

# Next
# - replace the variables with more intuitive information 
# - make a list of the census tracts of interest 
# - filter the tracts in the age / race df 
# - compare across the groups 
# for clarity: the estimate:total is the sum of the total male and the total female: could be used for total number of that ethnicity
```

