---
title: "Final Infographic"
author: "Luna Herschenfeld-Catalán she/her"
date:  2024-03-09
format:
  html:
    embed-resources: true
output:
  html_document:
    code_folding: hide
---

```{r include = FALSE}

# Read in libraries
library(tidyverse)
library(here)
library(lubridate)
library(janitor)
library(sf)
library(tidycensus)
library(patchwork)

# do google fonts
library(monochromeR)
library(showtext)
library(ggtext)
library(ggrepel)

knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

# Data

## Wrangle and Clean Data

https://www.census.gov/cgi-bin/geo/shapefiles/index.php


```{r message = FALSE, warning = FALSE, results='hide'}
# geometry of the puma 2022 shapefiles
puma <- read_sf(here("data", "tl_2022_06_puma20", "tl_2022_06_puma20.shp"))

puma_la <- puma %>% 
  filter(grepl(037, PUMACE20)) %>% 
  rename(PUMA = PUMACE20)

pums_housing_22 <- read_csv(here::here("data", "pums_2022.csv"))

# https://www.huduser.gov/portal/datasets/il/il2020/2020summary.odn
people <- c(1, 2, 3, 4, 5, 6, 7, 8)
very_low <- c(39450, 45050, 50700, 56300, 60850, 65350, 69850, 74350)
extreme_low <- c(23700, 27050, 30450, 33800, 36550, 39250, 41950, 44650)
low <- c(63100, 72100, 81100, 90100, 97350, 104550, 111750, 118950)

income_breaks <- data.frame(people, extreme_low, very_low, low)

# remove the duplicates from 2022 data
puma_house_codes <- pums_housing_22 %>% 
  select(SERIALNO, PUMA) %>% 
  distinct()

puma_22 <- pums_housing_22 %>% 
  inner_join(puma_la, by = "PUMA") %>% 
  select(SERIALNO, WGTP, NP, HINCP, FINCP, GRPIP) %>% 
  mutate(NP = as.numeric(NP),
         HINCP = as.numeric(HINCP),
         FINCP = as.numeric(FINCP),
         GRPIP = as.numeric(GRPIP)) %>% 
  group_by(SERIALNO) %>% 
  summarize_all(mean, na.rm = TRUE) %>% 
  left_join(puma_house_codes)

la_tract_vacant <- get_acs(
  state = "CA",
  county = "Los Angeles",
  geography = "tract",
  variables = "B25004_001",
  geometry = TRUE,
  year = 2022
)

la_tract <- la_tract_vacant %>%
  mutate(NAME = gsub("; Los Angeles County; California", # elements that you want to remove
                                  "", # replace with blank
                                  NAME)) %>%
  mutate(NAME = gsub("Census Tract ", # elements that you want to remove
                                  "", # replace with blank
                                  NAME)) %>% 
  filter(GEOID != "06037599100") %>% # islands
  filter(GEOID != "06037599000") %>% # islands
  filter(GEOID != "06037980003") %>%
  filter(GEOID != "06037980004") %>%
  filter(!(NAME >= 9000 & NAME <= 9800))
  
```

## Looking at the places where affordable housing projects started

```{r, message=FALSE}
# read in affordable housing projects in LA from 2003 to present
ah_raw <- read_csv(here("data/LAHD_Affordable_Housing_Projects_List__2003_to_Present__20240119.csv")) %>% 
  clean_names()

# clean affordable housing data 
ah_clean <- ah_raw %>% 
  mutate(fun_date = as.Date(date_funded, tryFormats = c("%m/%d/%Y"))) %>% # make date_funded as date
  mutate(year = lubridate::year(fun_date)) %>% # make into year column 
  select(name, year, fun_date, construction_type, site_community, total_units = project_total_units, 
         housing_type, lahd_funded, in_service_date, gps_coords_on_map) %>% 
  mutate(gps_coords_on_map = gsub("[POINT()]", # elements that you want to remove
                                  "", # replace with blank
                                  gps_coords_on_map)) %>% # remove these elements from gps column 
  separate_wider_delim(gps_coords_on_map, 
                       delim = " ", names = c("empty", "coords"), # separate space from before the coordinates
                       too_many = "merge") %>% 
  separate_wider_delim(coords, delim = " ", names = c("Longitude", "Latitude"), # split lat and long coords
                       too_many = "merge") %>% 
  select(-empty) %>% 
  st_as_sf(coords = c("Longitude", "Latitude"), # make into geometry object
                 crs = st_crs(la_tract)) %>% 
  st_join(la_tract) %>% 
  mutate(site_community = str_to_title(site_community))
```

# Assigning income categories to households

## Create income breaks for households above 8 (from 9-20)

```{r message = FALSE}
# very low income 
size <- seq(9, 20, 1)
very_low <- data.frame()
row_n <- data.frame()

for (i in seq_along(size)) {
  income <- 1.32 + (i*.08)
  limit <- income*56300
  row <- size[i]
  row_n <- rbind(row_n, row)
  very_low = rbind(very_low, limit)
}

very_low <- cbind(row_n, very_low) %>% 
  rename(people = X9,
         very_low = X78820)

extreme <- data.frame()
for (i in seq_along(size)) {
  # https://www.huduser.gov/portal/datasets/il/il2020/2020ILCalc3080.odn#calculator
  income <- 44120 + (i*4480)
  extreme <- rbind(extreme, income)
}

low <- data.frame()
for (i in seq_along(size)) {
  income <- 1.32 + (i*.08)
  limit <- income*90100
  low <- rbind(low, limit)
}

cols_9_20 <- cbind(extreme, very_low) %>% 
  rename(extreme_low = X48600) %>% 
  cbind(low) %>% 
  rename(low = X126140)

income_breaks <- rbind(income_breaks, cols_9_20)
```

## Creating For Loop

### Applying for loop to the whole data set

```{r message = FALSE}

puma_22$status <- rep(NA, nrow(puma_22))

for (person in 1:nrow(puma_22)) {
  
  individual <- puma_22$NP[person]
  
  income <- income_breaks %>% 
    filter(people == individual)
  
  #print(individual)
  #print(individual)
  # Define thresholds for labels
  extreme_threshold <- income[[2]]
  very_low_threshold <- income[[3]]
  low_threshold <- income[[4]]
  
  # # Assign labels based on thresholds
  puma_22$status[person] <- ifelse(puma_22[["HINCP"]][person] <= extreme_threshold, 
                              "extreme",
                              ifelse(puma_22[["HINCP"]][person] > extreme_threshold & puma_22[["HINCP"]][person] <= very_low_threshold, 
                                     "very low",
                                     ifelse(puma_22[["HINCP"]][person] > very_low_threshold & puma_22[["HINCP"]][person] <= low_threshold, 
                                            "low",
                                            ifelse(puma_22[["HINCP"]][person] > low_threshold, "not eligible", "non"))))
}

# uncount to get the total number of households in each census tract with each status
puma_wgt <- puma_22 %>% 
  uncount(WGTP)

```

# Mapping census tracts in LA

```{r message = FALSE}
puma_geo <- puma_la %>% 
  select(PUMA, geometry)


puma_plot <- puma_wgt %>% 
  group_by(PUMA) %>% 
  count(status) %>% 
  pivot_wider(names_from = "status", values_from = "n") %>% 
  ungroup() %>% 
  clean_names() %>% 
  select(puma, extreme, very_low, low, not_eligible) %>% 
  mutate(PUMA = puma, 
         total_households = rowSums(.[2:5]),
         total_eligible = rowSums(.[2:4]),
         percent_eligible = (total_eligible/total_households) * 100 ) %>% 
  left_join(puma_geo) %>% 
  st_as_sf()

# crop puma plot with the la_tract to remove the top part of LA
puma_plot_crop <- st_crop(puma_plot, la_tract)


# has 2196 points
tract_neighborhood <- read_csv(here("data/Census_Tract_Locations__LA__20240206.csv")) %>% 
  mutate(census_tract = gsub(", Los Angeles County, California", # elements that you want to remove
                                  "", # replace with blank
                                  Tract)) %>% 
  select(-Tract, census_tract, Neighborhood, Latitude, Longitude) %>% 
  st_as_sf(coords = c("Longitude", "Latitude"), # make into geometry object
                 crs = st_crs(puma_plot)) %>% 
  st_crop(puma_plot_crop) # neighborhoods only in the census tracts of interest

# make a df with the puma information and census tract geometry 
puma_tract <- puma_plot_crop %>% # has 67 observations
  st_join(tract_neighborhood,
                join = st_intersects) %>% # has the same number of points as the neighborhood
  clean_names() %>% 
  select(neighborhood, census_tract, location, puma, geometry) %>% 
  mutate(census_tract = gsub("Census Tract ", # elements that you want to remove
                                  "", # replace with blank
                                  census_tract))


# rename the tract column
la_tract <- la_tract %>% 
  rename(census_tract = NAME) %>% 
  select(census_tract, geometry)

puma_tract <- puma_tract %>% 
  st_drop_geometry() %>% 
  left_join(la_tract)

puma_eligibility <- puma_plot %>% 
  st_drop_geometry() %>% 
  select(!PUMA)

# make map of the census tracts with the puma household data
tract_eligibility <- puma_tract %>% 
  left_join(puma_eligibility, by = "puma") %>% 
  filter(!st_is_empty(geometry)) %>% 
  st_as_sf()
```

# Looking at census tract and the percent of income spent on rent: Rent Burdened

```{r}

rent_burden <- puma_22 %>% 
  group_by(PUMA) %>% 
  rename(percent_rent = GRPIP) %>%  
  select(WGTP, percent_rent, PUMA) %>% 
  mutate(status = ifelse(percent_rent >= 30, "burden", "not_burden")) %>% 
  uncount(WGTP) %>% # apply the weights
  count(status) %>% 
  pivot_wider(names_from = "status", values_from = "n") %>% 
  ungroup() %>% 
  rename(puma = PUMA) %>% 
  mutate(total = burden+not_burden,
         percent_burden = (burden/total)*100)
  
#plot(rent_burden_2$geometry)

rent_burdened <- puma_tract %>% 
  left_join(rent_burden) %>% 
  filter(!st_is_empty(geometry)) %>% 
  st_as_sf()

```

# Creating a custom theme

```{r}
library(monochromeR)
library(showtext)
library(ggtext)
library(ggrepel)

# import google fonts 
font_add_google(name = "Josefin Sans",
                family = "josefin") # name we provide ggplot

font_add_google(name = "Sen",
                family = "sen") # name we provide ggplot

font_add_google(name = "Tenor Sans",
                family = "tenorSans") # name we provide ggplot

# Pairs
font_add_google(name = "Playfair Display",
                family = "playfairDisplay") # name we provide ggplot

font_add_google(name = "Alice",
                family = "alice") # name we provide ggplot

# enable show text here that configures font across platforms
showtext_auto()

my_theme_dark <- theme(
  plot.title = element_text(size = 20, face = "bold", family = "playfairDisplay", hjust = 0.5), # Font size set to 16 and bold.
  plot.caption.position = "plot",
  plot.caption = element_text(size = 10, hjust = 0, family = "alice", color = "white"),
  axis.title = element_text(size = 14), # Font size set to 12.
  axis.text = element_blank(), # no axis text.
  axis.ticks = element_blank(),
  legend.title = element_text(size = 15, face = "bold", family = "alice", color = "white"), # Font size of the title of the legend set to 12 and bold.
  legend.text = element_text(size = 12, color = "white"), # font size of the text in the legend
  plot.background = element_rect(fill = "black"),
  panel.background = element_rect(fill = "black"),
  legend.background = element_rect(fill = "black"),
  panel.grid.major = element_blank(),
  plot.margin = margin(12, 6, 12, 18),
  panel.grid.minor = element_blank(), # invisible auxiliary grids
  strip.background = element_rect(fill = "black", color = "#ff975d"), # strip background
  strip.text = element_text(size = 12, face = "bold", family = "alice", color = "white"), # strip texts
  legend.position = "top", # position of the legend
  legend.box.background = element_rect(color = "black"), # background of the plot
  legend.key.size = unit(5, "cm") # size of the legends key
)
```

# Figures

==

```{r fig.width = 11, fig.height = 20}

my_brew_palette15 <- RColorBrewer::brewer.pal(n = 20, name = 'YlGnBu')
my_brew_palette10 <- RColorBrewer::brewer.pal(n = 10, name = 'RdBu')
my_brew_palette5 <- RColorBrewer::brewer.pal(n = 10, name = 'Reds')

# eligibility of housing in LA
tract_eligibility_percent <- ggplot(tract_eligibility) +
  geom_sf(aes(fill = percent_eligible)) +
  scale_fill_stepsn(colors = my_brew_palette5,
                    labels = scales::label_percent(scale = 1),
                    breaks = scales::breaks_width(width = 7)) +
  labs(title = "Exploring Affordable Housing Needs in Los Angeles County, CA",
       fill = "Percent Eligible",
       caption = "Households Eligible for Affordable Housing by PUMA in Los Angeles, CA") + 
  guides(fill = guide_colorbar(barheight = unit(1, units = "mm"),  
                                 barwidth = unit(100, units = "mm"),
                                 direction = "horizontal",
                                 ticks.colour = "grey20",
                                 title.position = "top",
                                 label.position = "bottom",
                                 title.hjust = 0.5)) +
  my_theme_dark




rent_burden_plot <- ggplot(rent_burdened) +
  geom_blank() +
  geom_sf(aes(fill = percent_burden)) +
  scale_fill_stepsn(colors = my_brew_palette5,
                    breaks = scales::breaks_width(width = 10),
                    labels = scales::label_percent(scale = 1)) +
  labs(title = "Exploring Affordable Housing Needs in Los Angeles County, CA",
       fill = "Percent Burdened",
       caption = "Percent of Households Rent Burdened by Census Tract in Los Angeles, CA (2022). Dots represent location of affordable housing projects from 2003-2020.") +
  guides(fill = guide_colorbar(barheight = unit(1, units = "mm"),  
                                 barwidth = unit(100, units = "mm"),
                                 direction = "horizontal",
                                 ticks.colour = "grey20",
                                 title.position = "top",
                                 label.position = "bottom",
                                 title.hjust = 0.5)) +
  my_theme_dark


burden_affordable_projects <- rent_burden_plot +
  geom_sf(data = ah_clean, alpha = 0.5, size = 1)

tract_eligibility_percent
burden_affordable_projects

```

# Selecting the Census Tracts of interest

```{r}

# joining the eligibility and rent data together by census tract
burden_eligible <- rent_burdened %>% 
  st_drop_geometry() %>% 
  left_join(tract_eligibility, by = "census_tract") %>% 
  st_as_sf()

top_burden <- burden_eligible %>% 
  slice_max(order_by = percent_burden,
            n = 200)
plot(top_burden$geometry)

# high burden, no affordable housing
burden_non <- top_burden %>% 
  slice_min(order_by = percent_burden, 
            n = 25) 

plot(burden_non $geometry)
# high burden, affordable housing
burden_ah <- burden_eligible %>% 
  slice_max(order_by = percent_burden,
            n = 30)

plot(burden_ah$geometry)

# low burden, with affordable housing - and also have high percentage eligible for affordable housing
lowburden_ah <- burden_eligible %>% 
  slice_max(order_by = percent_eligible,
            n = 450) %>% 
  slice_min(order_by = percent_eligible,
            n = 25)

plot(lowburden_ah$geometry)

# low burden, no affordable housing - low percentage of households that qualify for affordable housing
lowburden_non <- burden_eligible %>% 
  slice_min(order_by = percent_eligible,
            n = 100) %>% 
  slice_min(order_by = percent_eligible,
            n = 20)

plot(lowburden_non$geometry)

# neighborhoods right next to undeserved geometry represented by `burden_non`
interest <- burden_eligible %>% 
  slice_min(order_by = percent_burden,
            n = 600) %>% 
  slice_max(order_by = percent_burden,
            n = 15) 

plot(interest$geometry)
# %>% 
#   st_union()

```

# Plotting the three comparison neighborhoods

```{r fig.width = 11, fig.height = 11}

# high burden, no affordable housing
burden_non_plot <- burden_non %>% 
  st_union() %>% 
  ggplot() +
  geom_sf() +
  labs(title = "Neighborhoods with High Rent Burden, \nNo Affordable Housing") +
  my_theme_dark

# high burden, affordable housing
burden_ah_plot <- burden_ah %>% 
  st_union() %>% 
  ggplot() +
  geom_sf() +
  labs(title = "Neighborhoods with High Rent Burden, \nWith Affordale Housing") +
  my_theme_dark

# low burden, with affordable housing - and also have high percentage eligible for affordable housing
lowburden_ah_plot <- lowburden_ah %>% 
  st_union() %>% 
  ggplot() +
  geom_sf() +
  labs(title = "Neighborhoods with Low Rent Burden, \nWith Affordale Housing",
       subtitle = "and also have high percentage eligible for affordable housing") +
  my_theme_dark

# ow percentage of households that qualify for affordable housing
lowburden_non_plot <- lowburden_non %>% 
  st_union() %>% 
  ggplot() +
  geom_sf() +
  labs(title = "Neighborhoods with Low Rent Burden, \nNo Affordale Housing",
       subtitle = "low percentage of households that qualify for affordable housing") +
  my_theme_dark

interest_plot <- interest %>% 
  st_union() %>% 
  ggplot() +
  geom_sf() +
  labs(title = "Neighborhood with low rent burden right next \nto neighborhood with high rent burden and no support",
       subtitle = "50% percentage of households that qualify for affordable housing, less than 20% rent burden") +
  my_theme_dark

burden_non_plot / burden_ah_plot / lowburden_ah_plot / lowburden_non_plot / 
interest_plot

```

```{r}

unique(burden_non$neighborhood.x)

```

The neighborhoods are "Commerce", "Maywood", "Bell", "Bell Gardens", "Cudahy", "Huntington Park". 

```{r fig.width = 11, fig.height = 11}
# title = "Exploring Affordable Housing Needs in Los Angeles County, CA"
my_brew_palette1 <- RColorBrewer::brewer.pal(n = 9, name = 'YlOrBr')

#colorspace::hcl_palettes()
library(colorspace)
hcl_palettes(plot = TRUE)
colors <- sequential_hcl(9, palette = "Red-Yellow")
colors

base_map <- ggplot(rent_burdened) +
  geom_blank() +
  geom_sf(aes(fill = percent_burden), color = "black", lwd = 0.05) +
  scale_fill_stepsn(colors = rev(colors),
                    breaks = scales::breaks_width(width = 5),
                    labels = scales::label_percent(scale = 1)) +
  guides(fill = guide_colorbar(barheight = unit(1, units = "mm"),
                                 barwidth = unit(220, units = "mm"),
                                 direction = "horizontal",
                                 ticks.colour = "grey20",
                                 title.position = "top",
                                 label.position = "bottom",
                                 title.hjust = 0.5)) +
  geom_sf(data = ah_clean, aes(color = "skyblue"), alpha = 0.5, size = 1, color = "skyblue") +
  labs(fill = "Percent of Census Tract Burdened") +
  guides(color = "Hello") +
  my_theme_dark

base_map
```


```{r fig.width = 11, fig.height = 11}
# burden_ah points
burden_ah_pt <- data.frame(
  y = c(34.0597395),
  x = c(-118.982),
    yend = c(33.83291),
  xend = c(-118.66136)
  # yend = c(33.62815),
  # xend = c(- 118.30718)
)

lowburden_ah_pt <- data.frame(
  y = c(34.245882),
  x = c(-118.4209225),
  yend = c(34.31389),
  xend = c(-118.12637)
)

lowburden_non_pt <- data.frame(
  y = c(33.860177),
  x = c(-118.398831),
  yend = c(33.82235),
  xend = c(-118.53424)
)

# burden_non points
burden_non_pt <- data.frame(
  y = c(33.99),
  x = c(-118.13795),
  #     yend = c(33.93291),
  # xend = c(-118.66136)
  yend = c(33.877222),
  xend = c(-117.92694)
)
 
# add the lines to the plot in segments - still editing this 
base_map +
    # geom_sf(data = burden_non %>% st_union(),
    #       fill = "#A9565A",
    #       color = "white") +
    geom_curve(data = burden_non_pt,
             aes(x = x, y = y, xend = xend, yend = yend),
             arrow = arrow(length = unit(0.01, "npc"), type = "open"),
             #linetype = "dashed",
             curvature = -0.3,
             size = 1.5,
             color = "#AA592C")
    # geom_curve(data = burden_non_pt,
    #          aes(x = x, y = y, xend = xend, yend = yend),
    #          arrow = arrow(length = unit(0.01, "npc"), type = "open"),
    #          #linetype = "dashed",
    #          curvature = -0.3,
    #          size = 1.3,
    #          color = "white")

```


```{r fig.width = 6, fig.height = 6}

# Neighborhoods 
n_labels <- data.frame(long = c(33.968132, 33.99926595, 33.9924535, 33.9583695),
                     lat = c(-118.154518, -118.152339, -118.195177, -118.1809385), 
                     label = c("Bell Gardens", "Commerce", "Maywood", "Cudahy"))

ggplot(data = burden_non %>% st_union()) +
  geom_sf(fill = "#AA592C", color = "white", 
          lwd = 0.2) +
  geom_point(data = n_labels, 
             aes(x = lat, y = long),
             color = "white") +
  geom_text(data = n_labels, 
            aes(x = lat, y = long, label = label),
            vjust = -1, 
            color = "white", 
            family = "alice",
            size = 9) +
  #labs(caption = "Over 75% of the census tracts in the neighborhoods of \nCommerce, Maywood, Bell Gardens, and Cudahy are\n eligible for affordable housing in 2022. Over 40% of these \nlocations are also rent burdened, highlighting the need \nfor some sort of intervention. No affordable housing \ndevelopments have been built here in the last two \ndecades, indicating an area where affordable housing \nwould help alleviate rent burden for thousands of \nhouseholds.") +
  guides(fill = "none") +
  my_theme_dark +
  theme(plot.caption = element_text(size = 15)
        #plot.margin = margin(1, 300, 30, 30)
        )



```

```{r fig.width = 13, fig.height = 5}
library(gghighlight)
RColorBrewer::brewer.pal(n=10,"Reds")

order_2015 <- data.frame(
  y = c(210000000),
  x = c(2012),
  yend = c(50000000),
  xend = c(2015)
)

order_2022 <- data.frame(
  y = c(250000000),
  x = c(2024),
  yend = c(190000000),
  xend = c(2022)
)

# looking at the total construction of affordable housing in each year 
ah_clean_plot <- ah_clean %>% 
  drop_na() %>% 
  group_by(year) %>% 
  summarize(total_funding = sum(lahd_funded))

ah_clean_plot %>%  
  ggplot() +
  # geom_col(color = "#DDB472",
  #          fill = "black") +
  geom_col(data = ah_clean_plot %>% 
             filter(year != "2015"),
           aes(x = year, y = total_funding),
           alpha = 0.7,
           color = "#DDB472",
         fill = "black") +
  geom_col(data = ah_clean_plot %>% filter(year == 2015),
           aes(x = year, y = total_funding),
         color = "#DDB472",
         fill = "#DDB472") +
  geom_col(data = ah_clean_plot %>% filter(year == 2022),
           aes(x = year, y = total_funding),
         color = "#DDB472",
         fill = "#DDB472") +
  scale_y_continuous(expand=c(0,0),
                     labels = function(x) paste0(x / 1e6, "M")) +
  geom_curve(data = order_2015,
               aes(x = x, y = y, xend = xend, yend = yend),
              arrow = arrow(length = unit(0.01, "npc")),
             size = 0.7,
             curvature = -0.6,
             color = "white") +
    geom_curve(data = order_2022,
               aes(x = x, y = y, xend = xend, yend = yend),
              arrow = arrow(length = unit(0.01, "npc")),
             size = 0.7,
             curvature = 0.3,
             color = "white") +
  xlim(2000, 2035) +
  #gghighlight(year == 2015) +
  annotate(geom = "text", x = 2009, y = 250000000, 
           label = "Mayor Eric Garcetti passes \nExecutive Directive 13, \nSupport for Affordable Housing (ED 13)",
           color = "white",
           family = "alice",
           size = 6) +
  annotate(geom = "text", x = 2028, y = 250000000, 
           label = "Mayor Karen Bass passes \nExecutive Directive 1, \nExpedition of Permits (ED 1)",
           color = "white",
           family = "alice",
           size = 6) +
  theme_classic() +
  labs(x = 'Year',
       y = 'Total Funding (in millions)'
  ) +
  theme(plot.title = element_text(size = 20, face = "bold", family = "playfairDisplay", hjust = 0.5), # Font size set to 16 and bold.
        plot.caption.position = "plot",
        plot.caption = element_text(size = 10, hjust = 0, family = "alice", color = "white"),
        axis.title = element_text(size = 14), # Font size set to 12.
        axis.text = element_text(size = 16, color = "white", family = "alice"),
        #axis.ticks.x = element_line(size = 10, color = "white"),
        legend.title = element_text(size = 15, face = "bold", family = "alice", color = "white"), # Font size of the title of the legend set to 12 and bold.
        legend.text = element_text(size = 10, color = "white"), # font size of the text in the legend
        plot.background = element_rect(fill = "black"),
        panel.background = element_rect(fill = "black"),
        legend.background = element_rect(fill = "black"),
        plot.margin = margin(20, 6, 12, 20),
        axis.line.x = element_line(color = "white"),
        panel.grid.major.y = element_line(color = "white", size = 0.25), # invisible auxiliary grids
        strip.background = element_rect(fill = "black", color = "#ff975d"), # strip background
        strip.text = element_text(size = 12, face = "bold", family = "alice", color = "white"), # strip texts
        legend.position = "top", # position of the legend
        legend.box.background = element_rect(color = "black"), # background of the plot
        legend.key.size = unit(1, "cm") # size of the legends key
  )

# Mayor Karen Bass passes Executive Directive Order 1, Expedition of Permits (ED 1)

#lahd_funding
```

# Next steps

-   get information for these areas
-   Number of households (approximate)
-   How has that changed since 2003

# Look at the industry / jobs of people in the census tract

This data does not exist

```{r}
# load list of variables that I could access
variables <- load_variables(2022, "acs1", cache = TRUE)

#B24114 is for Detailed Occupation for the Civilian Employed Population 16 Years and Over
#B24115 is for Detailed Occupation for the Civilian Employed Population 16 Years and Over - Male
#B24116 is for Detailed Occupation for the Civilian Employed Population 16 Years and Over - Female
B24114 <- variables %>% 
  filter(str_detect(name, "B24114"))

#B24114$name

la_tract_occupation <- get_acs(
  state = "CA",
  county = "Los Angeles",
  geography = "state",
  variables = "B24114_100",
  geometry = TRUE,
  year = 2021
)

v <- data(acs5_geography)
```

## Try this that looks at the number of workers in household based on size

```{r}
# B08202 Household Size by Number of Workers in Household

B08202 <- variables %>% 
  filter(str_detect(name, "B08202"))

la_tract_workers <- get_acs(
  state = "CA",
  county = "Los Angeles",
  geography = "state",
  variables = B08202$name,
  geometry = TRUE,
  year = 2022
)
```

# Look at the average age / demographic data by tract

https://walker-data.com/tidycensus/articles/basic-usage.html

```{r}
ah_tracts <- unique(ah_clean$NAME)
```

```{r}
# B01001A is Sex by Age (White Alone)
# B01001B is Sex by Age (Black or African American Alone)
# B01001

 B01001 <- variables %>% 
  filter(str_detect(name, "B01001"))

# has the de
la_tract_age <- get_acs(
  state = "CA",
  county = "Los Angeles",
  geography = "tract",
  variables = B01001$name,
  geometry = TRUE,
  year = 2022
)

tract_ethnicity <- la_tract_age %>%
  mutate(NAME = gsub("; Los Angeles County; California", # elements that you want to remove
                                  "", # replace with blank
                                  NAME)) %>%
  mutate(NAME = gsub("Census Tract ", # elements that you want to remove
                                  "", # replace with blank
                                  NAME)) %>% 
  rename(census_tract = NAME)

#unique(tract_ethnicity$census_tract)

# - make a list of the census tracts of interest 
on_plot <- lowburden_ah %>% 
  mutate(area = "lowburden_ah") %>% 
  rbind(burden_ah %>% mutate(area = "burden_ah")) %>% 
  rbind(burden_non %>%  mutate(area = "burden_non")) %>% 
  rbind(lowburden_non %>% mutate(area = "lowburden_non")) %>% 
  select(census_tract, area)

# - filter the tracts in the age / race df 
sex_age_ethnicity <- tract_ethnicity %>% 
  
  # - replace the variables with more intuitive information 
  mutate(variable = gsub("B01001A",
                       "white",
                       variable),
         variable = gsub("B01001B",
                       "black",
                       variable),
         variable = gsub("B01001C",
                       "ai", # american indian and alaska native
                       variable),
         variable = gsub("B01001D",
                       "asian",
                       variable),
         variable = gsub("B01001E",
                       "haw", # hawaiian and pacific islander
                       variable),
         variable = gsub("B01001F",
                       "other",
                       variable),
         variable = gsub("B01001G",
                       "two", # two or more 
                       variable),
                  variable = gsub("B01001H",
                       "wn", # white non hispanic and latino
                       variable),
         variable = gsub("B01001I",
                       "hisp", # hispanic and latino
                       variable),
         variable = gsub("B01001_",
                       "total_",
                       variable)) %>% 
  separate_wider_delim(cols = 3,
                       delim = "_",
                       names = c("group", "age_variable")) 


#plot(ethnicity$geometry)
#unique(all_eth$census_tract)  
#unique(ah_tracts_eth$census_tract)

ah_tracts_eth <- sex_age_ethnicity %>% 
  filter(census_tract %in% ah_tracts) %>% 
  filter(age_variable == "001") %>% 
  select(-c(GEOID, age_variable, moe, geometry)) %>% 
  mutate(status = "affordable")
  # pivot_wider(names_from = group,
  #             values_from = estimate)

ah_tot <- ah_tracts_eth %>% 
  filter(group == "total") %>% 
  select(census_tract, group, estimate) %>% 
  rename(totals = group,
         total_est = estimate)

ah_tracts_eth <- ah_tracts_eth %>% 
  filter(group != "total") %>% 
  left_join(ah_tot, by = "census_tract") %>% 
  mutate(percent = round((estimate/total_est)*100, 1)) %>% 
  select(census_tract, group, estimate, percent, status) 

all_eth <- sex_age_ethnicity %>% 
  filter(census_tract %in% rent_burdened$census_tract) %>% 
  filter(age_variable == "001") %>% 
  select(-c(GEOID, age_variable, moe, geometry)) %>% 
  mutate(status = "non")

all_tot <- all_eth %>% 
  filter(group == "total") %>% 
  select(census_tract, group, estimate) %>% 
  rename(totals = group,
         total_est = estimate)

all_eth <- all_eth %>% 
  filter(group != "total") %>% 
  left_join(all_tot, by = "census_tract") %>% 
  mutate(percent = round((estimate/total_est)*100, 1)) %>% 
  select(census_tract, group, estimate, percent, status) 

# combine the tracts
compare <- all_eth %>% 
  rbind(ah_tracts_eth) %>% 
  rename(groups = group) %>% 
  filter(!groups %in% c("ai", "haw", "two"))


```

# Comparing the demographics of all of LA vs the tracts with affordable housing 

```{r fig.width = 8, fig.height = 11}
showtext_auto()
# Custom labels for x-axis
custom_labels <- c("White (non Hispanic \nor Latino)", "White", "Other", "Hispanic", "Black", "Asian")

subtitle <- "The demographics of the census tracts in all of the <span style = 'color: #fe9929; font-size: 11pt;'>**Los Angeles Metropolitan Area**</span> vs tracts that have<span style = 'color: #8c2d04; font-size: 11pt;'>**built affordable housing**</span> since 2003."

ggplot(compare) +
  geom_boxplot(aes(x = groups, y = percent, fill = status), 
               color = "white",
               size = 0.2) +
  scale_fill_manual(values = c("skyblue", "#DDB472")) +
  coord_flip() +
  labs(y = "Percent of Census Tract", 
       x = "",
      #subtitle = subtitle
       fill = ""
       ) +
  scale_x_discrete(labels = rev(custom_labels)) +
  theme_minimal() +
  guides(fill = "none") +
  theme(plot.subtitle = ggtext::element_textbox_simple(family = "alice", 
                                                       size = 10)) +
  theme(plot.title = element_text(size = 20, face = "bold", family = "playfairDisplay", hjust = 0.5), # Font size set to 16 and bold.
        plot.caption.position = "plot",
        plot.caption = element_text(size = 10, hjust = 0, family = "alice", color = "white"),
        axis.title = element_text(size = 14), # Font size set to 12.
        axis.text = element_text(size = 16, color = "white", family = "alice"),
        #axis.ticks.x = element_line(size = 10, color = "white"),
        legend.title = element_text(size = 15, face = "bold", family = "alice", color = "white"), # Font size of the title of the legend set to 12 and bold.
        legend.text = element_text(size = 10, color = "white"), # font size of the text in the legend
        plot.background = element_rect(fill = "black"),
        panel.background = element_rect(fill = "black"),
        legend.background = element_rect(fill = "black"),
        plot.margin = margin(12, 6, 12, 18),
        axis.line.x = element_line(color = "white"),
        panel.grid = element_blank(), # invisible auxiliary grids
        #panel.grid.minor.x = element_line(color = "white"),
        strip.background = element_rect(fill = "black", color = "#ff975d"), # strip background
        strip.text = element_text(size = 12, face = "bold", family = "alice", color = "white"), # strip texts
        legend.position = "top", # position of the legend
        legend.box.background = element_rect(color = "black"), # background of the plot
        legend.key.size = unit(1, "cm") # size of the legends key
  )


 colors
# subtitle = "The demographics of the census tracts in all of the <span style = 'color: #c19089; font-size: 20pt;'>**Los Angeles Metropolitan Area**</span> vs tracts that <span style = 'color: #67000D; font-size: 20pt;'>**have built affordable housing**</span> since 2003.")
```

```{r fig.width = 10, fig.height = 9}
showtext_auto()

custom_labels <- c("White (non Hispanic \nor Latino)", "White", "Other", "Hispanic", "Black", "Asian")

subtitle <- "Compared to the demographic distribution of <span style = 'color: #DDB472; font-size: 16pt;'>census tracts in Los Angeles overall</span>, the <span style = 'color: skyblue; font-size: 16pt;'>census tracts that have affordable housing</span> projects built since 2003 have a higher percentage of blank and hispanic individuals."
subtitle 

c_means <- compare %>% 
  group_by(status, groups) %>% 
  summarize(means = mean(percent, na.rm = TRUE),
            .groups = "drop")

# ggplot(compare %>% filter(groups == "white")) +
#   geom_density_ridges(aes(x = percent, y = groups, fill = status, linetype = status), 
#                       alpha = 0.5, 
#                       quantile_lines=TRUE,
#                       quantile_fun = mean) +
#   # geom_vline(data = c_means %>% filter(groups == "white"), aes(group = groups, 
#   #                                xintercept = means,
#   #                                linetype = status),
#   scale_y_discrete(expand = c(0.1, 0)) +
#   xlim(0,100)
library(ggridges)

ggplot(compare) +
  geom_density_ridges(aes(x = percent, y = groups, fill = status, linetype = status), 
                      alpha = 0.7, 
                      quantile_lines=TRUE,
                      quantile_fun = mean,
                      scale = 0.9) +
  scale_y_discrete(expand = c(0.01, 0.5),
                   labels = rev(custom_labels)) +
  xlim(0,100) +
  scale_fill_manual(values = c("skyblue", "#DDB472")) +
  geom_text(data = c_means, aes(x = means, y = groups, label = paste(round(means, 0)), color = status),
            family = "alice",
            vjust = 1, 
            hjust = 0,
            #nudge_x = -7
            size = 6) +
  scale_color_manual(values = c("skyblue", "#DDB472")) +
  labs(y = "Percent of Census Tract", 
       x = "Percent of Census Tract",
       #caption = subtitle,
       fill = ""
       ) +
  #scale_x_discrete(labels = rev(custom_labels)) +
  guides(fill = "none", linetype = "none", color = "none") +
  theme(plot.title = element_text(size = 20, family = "playfairDisplay", hjust = 0.5), # Font size set to 16 and bold.
        plot.caption.position = "plot",
        plot.title.position = "plot",
        plot.caption =  ggtext::element_textbox_simple(size = 15, hjust = 0, family = "alice", color = "white"),
        axis.title = element_text(size = 14), # Font size set to 12.
        axis.text.y = element_text(size = 16, color = "white", family = "alice"),
        axis.text.x = element_blank(),
        axis.title.x = element_text(size = 16, color = "white", family = "alice"),
        #axis.ticks.x = element_line(size = 10, color = "white"),
        legend.title = element_text(size = 15, face = "bold", family = "alice", color = "white"), # Font size of the title of the legend set to 12 and bold.
        legend.text = element_text(size = 10, color = "white"), # font size of the text in the legend
        plot.background = element_rect(fill = "black"),
        panel.background = element_rect(fill = "black"),
        legend.background = element_rect(fill = "black"),
        plot.margin = margin(12, 6, 12, 18),
        #axis.line.x = element_line(color = "white"),
        panel.grid = element_blank(), # invisible auxiliary grids
        panel.grid.minor.x = element_line(color = "white", size = 0.1, linetype = "dashed"),
        strip.background = element_rect(fill = "black", color = "#ff975d"), # strip background
        strip.text = element_text(size = 12, face = "bold", family = "alice", color = "white"), # strip texts
        legend.position = "top", # position of the legend
        legend.box.background = element_rect(color = "black"), # background of the plot
        legend.key.size = unit(1, "cm") # size of the legends key
  )
```


```{r eval = FALSE}

library(ggstream)

ggplot(compare) +
  geom_stream(aes(x = percent, y = status, fill = status), bw = 0.5, type = "proportional")) +
  scale_fill_manual(values = c("skyblue", "#DDB472"))

```



```{r eval = FALSE}
# for clarity: the estimate:total is the sum of the total male and the total female: could be used for total number of that ethnicity
ethnicity <- sex_age_ethnicity %>% 
  filter(census_tract %in% on_plot$census_tract) %>% 
  filter(age_variable == "001") %>% 
  select(-c(GEOID, age_variable, moe))

sum_on_plot <- on_plot %>% 
  left_join(ethnicity, by = "census_tract") %>% 
  select(-census_tract) %>%
  st_drop_geometry() %>% 
  group_by(area, group) %>% 
  summarize(estimate = sum(estimate))
  
ggplot() +
  geom_col(data = sum_on_plot, aes(x = area, y = estimate, fill = group),
           position = "dodge")
```

# Making shapefiles for the census tracts

```{r eval = FALSE}
# high burden, no affordable housing
# burden_non_shp
burden_non %>%
  st_drop_geometry() %>% 
  head(1) %>% 
  select(extreme, very_low, low, not_eligible, total_households, # sum of the other columns
         percent_eligible
         ) %>%
  mutate(p_ex = round((extreme/total_households)*100, 0),
         p_vl = round((very_low/total_households)*100, 0),
         p_l = round((low/total_households)*100, 0),
         p_non = round((not_eligible/total_households)*100, 0)) %>% 
  select(p_ex, p_vl, p_l, p_non) %>%
  pivot_longer(cols = 1:4,
               names_to = "status",
               values_to = "percent") %>%
  mutate(name = "low_burden") %>% 
  ggplot() +
  geom_bar(aes(x = 1, y = percent, fill = status), 
           stat = "identity", position = "stack") +
  scale_fill_viridis(discrete = TRUE) +
  theme_minimal() +
  labs(x = NULL, y = "Percentage", title = "Not Burdened")

# high burden, affordable housing
#buden_ah_shp <- 
burden_ah %>% 
  st_drop_geometry() %>% 
  head(1) %>% 
  select(extreme, very_low, low, not_eligible, total_households, # sum of the other columns
         percent_eligible
         ) %>%
  mutate(p_ex = round((extreme/total_households)*100, 0),
         p_vl = round((very_low/total_households)*100, 0),
         p_l = round((low/total_households)*100, 0),
         p_non = round((not_eligible/total_households)*100, 0)) %>% 
  select(p_ex, p_vl, p_l, p_non) %>%
  pivot_longer(cols = 1:4,
               names_to = "status",
               values_to = "percent") %>%
  mutate(name = "low_burden") %>% 
  ggplot() +
  geom_bar(aes(x = 1, y = percent, fill = status), 
           stat = "identity", position = "stack") +
  scale_fill_viridis(discrete = TRUE) +
  theme_minimal() +
  labs(x = NULL, y = "Percentage", title = "Burden with AH")

# low burden, with affordable housing - and also have high percentage eligible for affordable housing
#low_burden_shp <- 
library(viridis)
lowburden_ah %>% 
  st_drop_geometry() %>% 
  head(1) %>% 
  select(extreme, very_low, low, not_eligible, total_households, # sum of the other columns
         percent_eligible
         ) %>%
  mutate(p_ex = round((extreme/total_households)*100, 0),
         p_vl = round((very_low/total_households)*100, 0),
         p_l = round((low/total_households)*100, 0),
         p_non = round((not_eligible/total_households)*100, 0)) %>%
  select(p_ex, p_vl, p_l, p_non) %>% 
  pivot_longer(cols = 1:4,
               names_to = "status",
               values_to = "percent") %>%
  mutate(name = "low_burden",
         ypos = c(50, 95, 75, 17), # location of labels on the plot
         status = as.factor(status)) %>% 
  ggplot() +
  geom_bar(aes(x = 1, y = percent, fill = reorder(status, percent)), 
           stat = "identity", position = "stack") +
  # Add text
  geom_text(
    aes(1, y = ypos, label = reorder(status, percent)),
    hjust = 0,
    nudge_x = 0,
    colour = "white",
    size = 7,
    family = "alice"
  ) +
  scale_fill_viridis(discrete = TRUE) +
  labs(x = NULL, y = "Percentage", title = "Low Burden with AH") +
  plot_theme



        
```

```{r eval = FALSE}
# Load necessary libraries
library(ggplot2)
library(XML)
library(svglite)

# Example data
data <- data.frame(
  category = c("A", "B", "C"),
  value1 = c(10, 20, 30),
  value2 = c(15, 25, 35)
)

# Create SVG plot
svg("file:///Users/lunacatalan/Downloads/8z5bfP01.svg", width = 500, height = 500)

# Read SVG image outline
svg_image <- xmlParse("file:///Users/lunacatalan/Downloads/8z5bfP01.svg")

# Extract the image dimensions
svg_dims <- xmlAttrs(xmlRoot(svg_image))

# Create the data visualization on top of the SVG image
plot_gg<- ggplot(data, aes(x = category)) +
  geom_bar(aes(y = value1, fill = "Value1"), stat = "identity") +
  geom_bar(aes(y = value2, fill = "Value2"), stat = "identity") +
  scale_fill_manual(values = c("Value1" = "blue", "Value2" = "red")) +
  theme_minimal()

# Retrieve the plot
plot <- ggplot_gtable(ggplot_build(plot_gg))

# Embed the SVG image outline as the background
print(xmlSVG(svg_image), vp = viewport(width = as.numeric(svg_dims["width"]),
                                       height = as.numeric(svg_dims["height"]),
                                       xmax = 1, ymax = 1, 
                                       just = c("left", "bottom")))

# Print the data visualization on top of the SVG image
grid.draw(plot)

# End SVG plot
dev.off()

```
